---
title: "Can Income Raise Individual's Happiness?"
subtitle: "A Analysis of Income Rank and Well-Being"
author: 
  - Qingya Li
thanks: "Code and data are available at: https://github.com/QingyaLi/Can-Income-Raise-Individual-Happiness. The replication doi: https://doi.org/10.48152/ssrp-0kb2-y638"
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "The report replicates the paper “The Effects of Income Transparency on Well-Being: Evidence from a Natural Experiment” by Perez-Truglia (2020). The author estimates the correlation between income transparency and Norwegian individual subjective-well being. He builds difference in difference models by importing the German dataset and he finds the actual disclosure effect in Norway. He concludes the subjective well-being gap between the poor and the rich increases as the income transparency increases. The replicative results are similar to the original. We find individual income rank associates about 0.3 and 0.6 increase with their happiness and life satisfaction, respectively."
output:
  bookdown::pdf_document2
toc: true
bibliography: references.bib
---

```{r setup, include=FALSE}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(haven)
```

\newpage
# Introduction

Whether income can raise an individual’s happiness has been a controversial topic for many years. Economists and policymakers are interested in studying this topic due to its central importance. Many researchers believe in a significant relationship between individual income and individual happiness. For instance, the paper published by Clark, Frijters, and Shields(2008) studied the relative income term in the utility function, happiness, and utility and how does relative income affect other economic behaviors by conducting the models. They found that individual compares income to others or oneself [@relative]. Income is also positively correlated to individual happiness [@relative]. They concluded that we need to take relative income seriously. 

However, other researchers disagree that income does not play a significant role in well-being. According to the Esterlin Paradox, happiness does not rise as income continues to increase over time. By studying several countries' time series data, he summarized some conclusions: The United States case against the idea that national income can increase happiness [@paradox]. Also, even though Japan is a developed country with a high national income rank, its happiness rank is the third from the bottom among 20 countries [@paradox]. 

This report replicates the paper written by Perez-Truglia (2020). As tax records became accessible online in 2001, in Norway, people are permitted to review their income as well as any other people [@paper]. The author analyzes how does income transparency affects Norwegians' well-being. Perez-Truglia (2020) first create histograms and line charts to compare the google search volumes of each keyword in Norway. He finds Norwegians are interested in searching tax lists, typically when the tax agency realizes the income data. Then he conducts difference in difference models. To limit the time effect, he uses German data as a counterfactual, which does not have disclosure in 2001 [@paper]. By comparing the subjective well-being between Norway and Germany, Perez-Truglia (2020) can find the real discolored effect on subjective well-being in Norway. The respondent variables in the models are happiness and life satisfaction , respectively. He purposes that higher-income transparency increases 29% of the happiness gap between poor and rich individuals, and 21% life satisfaction gap. His conclusion is relative to the study by Clark, Frijters, and Shields (2008). People compare their income to others and which may lead to the rich well-being becoming higher and higher, the poor situation is converse. 

We replicate the original plots and models by using the analytic dataset from the paper. We estimate that the respondents’ income rank approximately raise 0.3 and 0.6 their happiness and life satisfaction, respectively (p values are less than 0.001). Meaning higher individual income rank is associated with higher individual happiness. Additionally, other factors such as gender, education, age can also affect subjective well-being. 

# Data

The report is written by the **R** statistical language [@citer], the packages of **tidyverse** [@tidyverse] and **dplyr** [@dplyr]. Stata files are read by **haven** packages [@haven]. In the report, we use **RMarkdom** [@rmarkdown] and **bookdown** [@bookdown] to output the results. The graphs are created by **ggplot2** [@ggplot2].

The original paper ^[The original paper is available at: https://doi.org/10.1257/aer.20160256] uses multiple sources which are in .dta format [@paper]. The codes are run by Stata. The study is mainly based on the dataset of Norwegian Monitor Survey, google searches by country, google searches over time, name browsing skattelister, and GSOEP. The GSOEP dataset is available at the German Socio-Economic Panel ^[The GSOEP dataset is available at: https://www.eui.eu/Research/Library/ResearchGuides/Economics/Statistics/DataPortal/GSOEP] which you need to require permission to access. 

The Norwegian Monitor data is collected through a cross-sectional survey (the survey is self-response by mail) by the research institute Ipsos MMI [@paper]. Thus, the population is Norwegians and the sample is the one who completes the survey. However, non-response bias may exist. The survey questions focus on Norwegians’ subjective well-being, which is measured by happiness and life satisfaction [@paper]. The dataset contains the information from every other year in 1985-2013 [@paper]. The respondence date is not recorded but the researchers consider respondents finished the survey during September and December [@paper]. As Perez-Truglia (2020) mentioned, the tax data for the previous year is released in October by the agency. Therefore, the respondents complete the survey when the income transparency is most significant, which may overestimate the effect of subjective well-being. Meanwhile, the survey in 2001 may finish before the tax record is accessible online which may lead to an underestimation [@paper]. There are 49,009 observations in the dataset are used in my report.

The google searches by country and the google searches over time data are collected by Google Trends, the datasets show the popularity of income searches in 2010[@paper]. Overall, four keywords are skatterlister, youtube, weather, and tax. Specifically, skatterlister means “tax list”, it refers to skattlister and skattelistene [@paper]. We will normalize the search keywords to a fraction of Youtube [@paper] and compare each search volumes in Figure \@ref(fig:fig1), which contains a robustness check by using Sweden data. We also uses line chart to show the number of Google search for each categories of keywords in Norway between 2010 and 2011 in \@ref(fig:fig2). 

Figure \@ref(fig:fig1) illustrates each of the annual searches volumes, relative to Youtube, between Norway and Sewden

```{r load data, echo = FALSE, message = FALSE, warning = FALSE}
google_searches_by_country <- read_dta("~/Downloads/starter_folder-main/inputs/data/111682-V2/google_searches_by_country.dta")
```

```{r fig1, fig.cap = 'The annual numbre of Google search for different categories of keywords, relative to the search for Youtube', fig.width = 10, fig.height = 6, warning = FALSE, message = FALSE, echo = FALSE}
google_searches_by_country <- read_dta("~/Downloads/starter_folder-main/inputs/data/111682-V2/google_searches_by_country.dta")
google_searches_by_country %>% 
  ggplot(mapping = aes(x=keyword, y= google))+
  geom_col(color = "dodgerblue", fill = "dodgerblue", alpha = 0.7)+
  theme_minimal()+
  facet_wrap(~country)+
  labs(x = "Search keyword", y="Google search relative to Youtube in %", title = "Annual search volumes, Norway vs. Sweden")

```

The bar plot shows that Norwegians are more likely to search weather than tax, and the times they search tax lists are much more than taxes. The evidence proves the uses of tax lists may not correlate to the taxes. Compare to Swedes, Norwegians are more interested in searching taxes but less interested in the weather. Overall, the search volumes for weather and Youtube are most likely between the two countries.

\newpage
Figure \@ref(fig:fig2) demonstrates the trend of the number of Google search for the keywords in Norway in 2010-2011.

```{r fig2,  fig.cap = 'The number of Google search for each categories of keywords in Norway from Jan, 2010 - Jan, 2011', fig.width = 10, fig.height = 6, warning = FALSE, message = FALSE, echo = FALSE}
google_searches_over_time <- read_dta("~/Downloads/starter_folder-main/inputs/data/111682-V2/google_searches_over_time.dta")
ggplot(google_searches_over_time, aes(x=edate)) + 
  theme_minimal()+
  geom_line(aes(y = skattelister, color = "skattelister"))+
  geom_line(aes(y = tax, color = "tax"))+
  geom_line(aes(y = weather, color = "weather"))+
  geom_line(aes(y = youtube, color = "youtube" ))+
  labs(x="Google search in %", y="Month", title = "Search volumes in 2010, Norway")
```


The google search for the tax lists is almost twice the search for taxes, and the search for weather is similar to the search for taxes. We notice that the search for tax lists suddenly raise a lot during October and even more than the search for Youtube. As we mentioned above, the tax agency released the tax data of the previous year every October which may cause an increase in searching tax lists. The evidence proves Norwegians are more interested in tax lists than other events.

# Model

According to the original paper [@paper], we intend to find the factors that may affect the subjective well-being of people by conducting difference in difference models. We predict that higher income transparency is associated with higher subjective well-being.

The replication baseline models are presented as below:

$$(1) SWB_{i,t} = \alpha_1 IncomeRank_{i,t} + \alpha_2 IncomeRank_{i,t}I^{01-13}_t + X_{i,t}\beta+\delta_t+\epsilon_{i,t}$$


- $SWB_{i,t}$ is subjective well-being measurment, where the higher scores of $SWB$ meaning higher well-being that individual feels
- $IncomeRank_{i,t}$ measures the individual's position in the national distribution of household income in a year, from 0 to 1. 0 represents individual gain the lowest income in the distribution
- $I^{01-13}_t$ is a dummy variable of income transparency. 1 is high income transparency, 0 is low, at the period of 2001-2013
- $X_{i,t}$ are the controls variables 
- $\delta_t$ is effects by years
- $\epsilon_{i,t}$ is random error term
- $i$ represents each individual and $t$ represents the year

$$(2) SWB_{i,t} = \alpha_1 IncomeRank_{i,t} + \alpha_2 IncomeRank_{i,t}I^{01-13}_t + \gamma IncomeRank_{i,t}(t-1985)+ X_{i,t}\beta+\delta_t+\epsilon_{i,t}$$ 


- $\gamma$ is the linear trend of the income and happiness between 1985 and 2013.


$$(3) SWB_{i,t} = \alpha_1 IncomeRank_{i,t} + \alpha_2  IncomeRank_{i,t}I^{01-13}_t 
+ \alpha_3  IncomeRank_{i,t}I^{97-00}_t+ X_{i,t}\beta+\delta_t+\epsilon_{i,t}$$


- $I^{97-00}$ is a dummy variable, it is equal to 1 if at the period of 1997-2000, 0 otherwise. It indicates the income rank just before the agency released income transparency. Assume the income transparency significantly increase the subjective well-being, we estimate that the $\alpha_2$ is positive and $\alpha_3$ is 0 meaning the period 1997-2000 does not affect the subjective well-being.


$$(4)SWB_{i,t} = \alpha_1 IncomeRank_{i,t}+\alpha_2IncomeRank_{i,t}I^{01-13}_t+ \alpha_3 HigherInternet_{i,t}$$

$$+\alpha_4 HigherInternet_{i,t}I^{01-13}_t +\alpha_5IncomeRank_{i,t} HigherInternet_{i,t}$$

$$+\alpha_6IncomeRank_{i,t}HigherInternet_{i,t}I^{01-13}_t + X_{i,t}\beta+\delta_t+\epsilon_{i,t} $$ 

- $HigherInternet_{i,t}$ is a dummy variable, it is 1 if individual has above-median internet access at home.The original paper use the triple-difference specification to improve the model [@paper]. The author consider there may exist factors other than income transparency will affect the subjective well-being [@paper]. Therefore, he creates this variable from internet access data to control the effects.

# Results

In order to replicate the results, we decide to run Ricardo's codes to generate the analytic dataset in **Stata** (the analytic dataset named "norway_data_ready_to_be_analyzed"), instead of using the raw Norwegian Monitor Survey dataset.
```{r import data, echo = FALSE, message = FALSE, warning = FALSE}
norway_data_ready_to_be_analyzed <- read_dta("~/Downloads/starter_folder-main/inputs/data/111682-V2/norway_data_ready_to_be_analyzed.dta")
```

```{r, include = FALSE, message = FALSE, warning = FALSE}
mod1 <- lm(po_happy ~ imp_hh_rank + post_2001 + post_2001*imp_hh_rank + year + marital_status + education + hh_size + hh_workers + female + age + age_sq,data = norway_data_ready_to_be_analyzed )
summary(mod1)
```

Table 1: Summary statistics in mod1

Variable | Estimate | P_value
---------|-----------|------------------
(Intercept)  |  `1.917e+00`  | `0.404691`  
imp_hh_rank  | `3.173e-01` | `< 2e-16`
post_2001  | `-7.563e-03` | `0.744517` 
year | `-5.822e-04` | `0.614883` 
marital_status |`-1.332e-01`  | `< 2e-16 `
education | `2.328e-02`  | `7e-06` 
hh_size | `6.620e-03`  | `0.142341` 
hh_workers  | `2.315e-02` |  `0.000184`
female | `1.157e-01`  | `< 2e-16` 
age  | `-3.737e-02` | < `2e-16` 
age_sq | `3.569e-04`  | `< 2e-16`
imp_hh_rank:post_2001 | `1.024e-01` | `0.000893` 


By applying the first model, we observe that the income rank is 0.31 positively correlated to happiness. P-value of income rank is less than 0.001, we have very strong evidence against the null hypothesis that there is no difference in the model with/without income rank. Besides, we notice that marital status and age are negatively correlated to happiness, while education, number of workers in the household, female, age square, the interaction between income rank and $I{2001–2013}$ are positively affecting happiness.


```{r, include = FALSE, message = FALSE, warning = FALSE}
mod2 <- lm(po_happy~ imp_hh_rank + post_2001 + post_2001*imp_hh_rank+trend*imp_hh_rank + marital_status+ education+hh_size+hh_workers+female+age+age_sq + year, data = norway_data_ready_to_be_analyzed)
summary(mod2)
```

Table 2: Summary statistics in mod2

Variable | Estimate | P_value
---------|-----------|------------------
(Intercept)  |  `7.593e-01`  | `< 2e-16`  
imp_hh_rank  | `3.211e-01 ` | `< 2e-16`
post_2001  | `-1.095e-02` | `0.7446 ` 
trend | `-3.420e-04` | `0.8691` 
marital_status |`-1.333e-01`  | `< 2e-16`
education | `2.330e-02`  | `6.95e-06` 
hh_size | `6.626e-03`  | `0.1420` 
hh_workers  | `2.328e-02` |  `0.0002`
female | `1.157e-01`  | `< 2e-16` 
age  | `-3.737e-02` | `< 2e-16`
age_sq | ` 3.569e-04 `  | `< 2e-16`
year | `NA`  | `NA`
imp_hh_rank:post_2001 | `1.093e-01` | `0.0603`
imp_hh_rank:trend | `-4.998e-04` | `0.8891`

We add the interaction of income rank with the period of 2001-2013 and the interaction of income rank with the trend. In this case, we find income rank is 0.32 positively significantly associated with happiness. 

```{r, include = FALSE, message = FALSE, warning = FALSE}
mod3 <- lm(po_happy~ imp_hh_rank + post_2001 + post_2001*imp_hh_rank+pre_2001 + pre_2001*imp_hh_rank + year+marital_status+ education+hh_size+hh_workers+female+age+age_sq, data = norway_data_ready_to_be_analyzed)
summary(mod3)
```

Table 3: Summary statistics in mod3

Variable | Estimate | P_value
---------|-----------|------------------
(Intercept)  |  `3.165e+00 `  | `0.239750 `  
imp_hh_rank  | `3.094e-01` | `< 2e-16`
post_2001  | `2.460e-03` | `0.933411` 
pre_2001  | `3.098e-03` | `0.914225`
marital_status |`-1.333e-01`  | `< 2e-16 `
education | `2.334e-02`  | `6.68e-06 ` 
hh_size | ` 6.709e-03`  | `0.137173` 
hh_workers  | `2.221e-02` |  ` 0.000387`
female | `1.155e-01`  | ` < 2e-16` 
age  | `-3.744e-02` | ` < 2e-16` 
age_sq | `3.576e-04`  | ` < 2e-16`
year | `-1.208e-03`  | `0.372410`
imp_hh_rank:post_2001 | `1.117e-01` | `0.001514`
imp_hh_rank:pre_2001 | `2.471e-02` | `0.598603`

We keep adding the interaction of income rank with at the period of 1985-2000 (before the tax agency released the income data). We have strong evidence that interaction between income rank and post period is 0.11 positively significantly associated with happiness (p-value is large than 0.001 but less than 0.05). 

```{r, include = FALSE, message = FALSE, warning = FALSE}
mod4 <- lm(po_happy~ imp_hh_rank + post_2001 + higher_internet + imp_hh_rank*higher_internet + post_2001*higher_internet + post_2001*imp_hh_rank+post_2001*imp_hh_rank*higher_internet+year+marital_status+education+hh_size+hh_workers+female+age+age_sq, data = norway_data_ready_to_be_analyzed)
summary(mod4)
```

Table 4: Summary statistics in mod4

Variable | Estimate | P_value
---------|-----------|------------------
(Intercept)  |  `6.971e-01`  | ` 0.763210`  
imp_hh_rank  | `3.507e-01` | `< 2e-16`
post_2001  | `3.127e-02` | `0.225897 ` 
higher_internet | `1.245e-01 ` | `0.000132` 
marital_status |`-1.323e-01`  | `< 2e-16 `
education | `1.752e-02 `  | `0.001559` 
hh_size | `2.003e-03`  | ` 0.670392 ` 
hh_workers  | `2.049e-02` |  `0.001325`
female | `1.214e-01`  | `< 2e-16` 
age | `-3.665e-02`  | `< 2e-16`
age_sq | ` 3.526e-04  `  | `< 2e-16`
year | `1.392e-05`  | `0.990457`
imp_hh_rank:higher_internet | `-1.280e-01` | `0.014546`
post_2001:higher_internet | `-1.736e-01` | `2.65e-05`
imp_hh_rank:post_2001 | `1.125e-02` | `0.818059`
imp_hh_rank:post_2001:higher_internet | `2.670e-01` | `0.000155`

The triple-difference specification model states higher internet is 0.12 significantly increased the happiness, interaction of income rank with post-2001 and higher internet is also raise 0.27 of happiness. In contrast, the interaction of income rank with higher internet, the interaction of post-2001 with higher internet low down happiness gradient.

As we mentioned above, subjective well-being is measured by both happiness and life satisfaction. Hence, we also use life satisfaction as a responding variable and apply the models. The results are included in Appendix. Both of the models support income rank raises life satisfaction (p-values are less than 0.001).

# Discussion

As GDP grows rapidly in many countries, overall national income raises. Economists argue that whether income can affect individual subjective well-being. The original paper “The Effects of Income Transparency on Well-Being: Evidence from a Natural Experiment” analyze the relationship between Norwegian income transparency and individual well-being since the residents are allowed to access the income information online since 2001 [@paper]. We intend to replicate the paper using the analytic datasets which can be generated by the provided codes. Similar to the original paper, we want to study the relationship between income rank and well-being (both happiness and life satisfaction). We first plot the histogram to show the Google search volumes of the four keywords: Youtube, weather, tax, and tax lists in Norway. We can observe that Norwegians are much more likely to search tax lists, meaning they are interested in the tax search tool. Also, the search of tax lists increases sharply when the agency released the income data. Then we build the models followed by the original paper. To control the time effect and other biases, adding a triple-difference specification and the variable of the pre-experimental year are necessary. Unlike the original paper also compare the results between Norway and Germany, we excluded this part. For some reason, we are unable to access the Germany dataset. So we only conduct the models using Norway dataset. 

We find our estimations prove the Easterlin Paradox which was demonstrated by economists Richard Easterlin. It shows that happiness is directly relatives to income within nations, while higher income is associated with higher happiness. Likewise, the relative income (RI) hypothesis also explains the relationship [@hypothesis]. The hypothesis shows that individual compares their income to others, and happiness depends on individual relative income rank [@hypothesis]. Many evidence supports the hypothesis. For instance, the researchers estimate the UK household dataset and conduct a longitudinal study [@comparison]. They use life satisfaction and General Health Questionnaire (GHQ) as dependent variables to measure utility proxies. They find the RI positively and significantly correlated to the two dependent variables in the spatial (region, occupation, industry) group [@comparison]. The results help economists and policy  makers realize individual income plays an essential role in individual well-being. It is necessary to for us develop national income and GDP.

In our report, some weaknesses should be improved in the future. First, the survey is sent by mail which asks respondents to complete several questions [@paper]. But non-respond bias exists. For example, some potential respondents may decide not to complete the survey form because they think they are not very happy. Second, we are supposed to set Norway data as a treatment group, Germany data as a control group in the difference in difference models. Then we compare the pre and post-period of individual well-being. It is better to do so since it can reduce the time effect and bias, also help us predict the actual policy effect. Third, there are many specifications besides higher internet (the internet access at home) that may affect subjective well-being. Thus, we should consider including more specifications in future studies. 


\newpage

\appendix

# Appendix {-}


```{r, include = FALSE, message = FALSE, warning = FALSE}
mod5<-lm(po_life_sat ~ imp_hh_rank + post_2001 +post_2001*imp_hh_rank +year + marital_status+ education + hh_size + hh_workers +female+ age+age_sq, data= norway_data_ready_to_be_analyzed )
summary(mod5)
```

Table 5: Summary statistics in mod5

Variable | Estimate | P_value
---------|-----------|------------------
(Intercept)  |  `-1.208e+01`  | `9.56e-05`  
imp_hh_rank  | `6.295e-01` | `< 2e-16`
post_2001  | `2.848e-02` | `0.4204` 
year | `6.137e-03 ` | `7.44e-05` 
marital_status |`-6.958e-02`  | `< 2e-16`
education | `4.820e-02 `  | `2.65e-12` 
hh_size | `-5.579e-03`  | `0.3424 ` 
hh_workers  | `-1.061e-02 ` |  `0.2493`
female | `1.531e-01`  | `< 2e-16 ` 
age  | `-3.702e-02` | `< 2e-16 `
age_sq | `-3.702e-02`  | `< 2e-16 `
imp_hh_rank:post_2001 | `1.142e-01` | `0.0475`


```{r, include = FALSE, message = FALSE, warning = FALSE}
mod6 <- lm(po_life_sat~ imp_hh_rank + post_2001 + higher_internet + imp_hh_rank*higher_internet + post_2001*higher_internet + post_2001*imp_hh_rank+post_2001*imp_hh_rank*higher_internet+year+marital_status+education+hh_size+hh_workers+female+age+age_sq, data = norway_data_ready_to_be_analyzed)
summary(mod6)
```

Table 6: Summary statistics in mod6

Variable | Estimate | P_value
---------|-----------|------------------
(Intercept)  |  `-1.336e+01`  | `1.76e-05`  
imp_hh_rank  | `5.674e-01` | `3.00e-10`
post_2001  | `5.142e-02` | `0.2154 ` 
higher_internet |` 3.540e-02`|  ` 0.6441 `                 
year | `6.788e-03` | `1.31e-05` 
marital_status |`-7.323e-02`  | `< 2e-16 `
education | `4.769e-02`  | `1.69e-11` 
hh_size | `-1.151e-02`  | `0.0595 ` 
hh_workers  | `-1.197e-02` |  `0.2166`
female | `1.576e-01`  | `< 2e-16` 
age  | `-3.690e-02` | `< 2e-16`
age_sq | ` 4.193e-04`  | `< 2e-16`
imp_hh_rank:higher_internet |`4.116e-02`|`0.7484 `
post_2001:higher_internet | `-1.081e-01` | `0.1765 `
imp_hh_rank:post_2001 | `5.000e-02` | ` 0.5928`
imp_hh_rank:post_2001:higher_internet | `1.784e-01` | `0.1910` 

\newpage


# References


